On the presentation of orthogonal groups over polynomial rings

### 2019-08-13 dl-paper.tex
Added V_T, finished proof of main result

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, R)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, R) &  } \] 
We denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of \[\St(\Phi, R) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times\] consisting of those triples $(g, h, u)$ for which
\[p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}\] belongs to $\overline{G}_M^{\geq 0}$.

@@

We denote by $\overline{V}$ the set of orbits of the action of $G_M^0$.
We use the notation $(g, [h], u)$ for the elements of $\overline{V}$.
We denote by $\overline{p}$ the function $\overline{V} \to \overline{G}^{\geq 0}_M/G_M^0$ sending each $(g, [h], u) \in V$ to the left coset $p(g, h, u)G_M^0$.

@@

We are ready to formulate the main result of this subsection.

\begin{prop} \label{prop43}
For $\Phi$ as in the statement of~\cref{lemma33}.
the operators $T_\alpha$ defined above specify a well-defined action of $\St(\Phi, B)$ on $\overline{V}$.

This action satisfies the following additional property. If $v = (g, [h], u) \in \overline{V}$ is such that $\overline{p}(v) = \overline{1} = G_M^0$ then for any $h_1 \in \St(\Phi, A[X\inv])$ holds $j_B(h_1) \cdot v = (h_1 \cdot g, [h \cdot h_1^{-1}], u)$ and, moreover, $\overline{p}(j_B(h_1) \cdot v) = \overline{1}$. Here $j_B$ denotes the obvious map $\St(\Phi, A[X\inv]) \to \St(\Phi, B)$.
\end{prop}
\begin{proof}
  By~\cref{superfluous-relations} and Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1} the action of $\St^{\leq 1}(\Phi, B)$ on $\overline{V}$ given by~\eqref{T_1} and~\eqref{T_leq0} is well-defined. On the other hand, the group $\St^{\leq 1}(\Phi, B)$ is isomorphic to $\St(\Phi, B)$ by~\cref{lemma33} (set $t = X^{-1}$).

@@

\subsection{Proof of the $\mathbb{P}^1$-glueing theorem.}
We retain notation of the previous subsection and assume additionally that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, R)$ is injective. This assumption allows us to consider $\St(\Phi, B)$ as a subgroup of $\St(\Phi, R)$.
%Denote by $I$ the ideal $M[X, X\inv]$ of $B$ (it is also an ideal of $R$).
Notice that by~\cref{lem:lemma32} the subgroup $\overline{G}^{\geq 0}_M$ is contained in $\St(\Phi, B)$.

%We denote by $t$ the map $\St(\Phi, R, I) \to \St(\Phi, B)$ from.
\begin{lemma} For $v\in \overline{V}$ and any $g \in \overline{G}^{\geq 0}_M$ holds $\overline{p}(g \cdot v) = g \cdot \overline{p}(v)$. \end{lemma}
\begin{proof} First of all, notice that the assertion of the lemma can be verified directly in the case $g = z_\alpha(f, a)$ for all $a\in A$, $f\in M[X]$. Using~\cref{SR:additivity} and the fact that $x_\beta(X^2f) \in P_\alpha(0)$ for all $\beta \in \Phi$ it is not hard to show that the assertion also holds for $g = z_\alpha(X^2f, aX^{-1})$.

It suffices to verify the assertion of the lemma for the generators $g = z_\alpha(f, \xi)$ of $\overline{G}_M^{\geq 0}$ for all $f \in M[X]$, $\xi\in A[X]$, $\alpha \in \Phi$. We accomplish this by 
induction on the degree of $\xi$ in $X$. The base of induction is clear.

Suppose that the assertion holds for all $g = z_\alpha(f, \xi)$ for which $\xi$ has degree $\leq n$.
Substituting $s := fX$, $\eta := X^{-1}$, $\xi := X\xi$ in~\eqref{Z5} we obtain the following equality in $\St(\Phi, B)$:
\begin{multline*}
 z_{\alpha+\beta}(f, X\xi) = x_\alpha(\epsilon Xf) \cdot x_{-\beta}(-faX^2) \cdot x_{\beta}(af) \cdot x_{\alpha+\beta}(f) \cdot \\ \cdot z_\alpha(-\epsilon Xf, -\epsilon \xi) \cdot x_{-\alpha}(-\epsilon \xi^2fX) \cdot x_{-\alpha-\beta}(- f \xi^2X^2) \cdot z_{-\beta}(X^2f\xi, -X^{-1}),\text{ where $\epsilon = N_{\alpha, \beta}$.}
\end{multline*}
From the first part of the proof and inductive assumption we obtain that the assertion of the lemma holds for all the factors in the right-hand side and, therefore, also holds holds for $g = z_\alpha(f, X\xi)$. It is easy to deduce from this that the assertion also holds for $g = z_\alpha(f, X\xi + a) = x_\alpha(-a) \cdot z_\alpha(f, X\xi) \cdot x_\alpha(a)$. 
\end{proof}

\begin{theorem}
 Let $\Phi$ be a root system of type $\rA_{\geq 4}, \rD_{\geq 5}$ or $\rE_{6,7,8}$.
 %Suppose that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, A[X, X\inv])$ is injective.  
 Then the map $j_-$ is injective and the following commutative square is pullback
 \begin{equation} \nonumber
  \xymatrix{ \St(\Phi, A) \ar[r] \ar[d] & \St(\Phi, A[X]) \ar[d]^{j_+} \\ \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X\inv]).}
 \end{equation}
\end{theorem}
\begin{proof}
 Since $j_- = j_R j_B$, where $j_B$ is as in~\cref{prop43}, it suffices to show that $j_B$ is injective.
 Set $v_0 = (1, [1], 1) \in \overline{V}$ and suppose $g \in \Ker(j_B)$.
 By~\cref{prop43} holds $v_0 = j_B(g) \cdot v_0 = (g, [g^{-1}], u)$ therefore $g \in G_M^0$ and consequently $g = 1$.

 Now suppose $g_+ \in \St(\Phi, A[X])$ and $g_- \in \St(\Phi, A[X\inv])$ are such that $j_-(g_-) = j_+(g_-)$. From~\cref{field-injectivity} it follows that the image $j_+(g_+) = j_-(g_-)$ in $\St(\Phi, k[X, X\inv])$ belongs to $\St(\Phi, k)$ and therefore coincides with the image in $\St(\Phi, k)$ of some $g_0 \in \St(\Phi, A)$. 
 Set $g_+' = g_+g_0^{-1}$ and $g_-' = g_-g_0^{-1}$.
 %Clearly, $g_+' \in \overline{\St}(\Phi, A[X], M[X])$. Set $v = () $
\end{proof}

>>>>

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, R)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, R) &  } \] 
We denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$.

Denote by $V$ the subset of \[\St(\Phi, R) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times\] consisting of those triples $(g, h, u)$ for which
\[p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}\] belongs to $\overline{G}_M^{\geq 0}$.

@@

We denote by $\overline{V}$ the set of orbits of the action of $G_M^0$.
We use the notation $(g, [h], u)$ for the elements of $\overline{V}$.
We denote by $\overline{p}$ the function $\overline{V} \to \overline{G}^{\geq 0}_M/G_M^0$ sending each $(g, [h], u) \in V$ to the left coset $p(g, h, u)G_M^0$.
The set $\overline{V}$ is the one upon which we will construct an action of $\St(\Phi, B)$.

Before we proceed further let us explain how our sets $V$ and $\overline{V}$ compare with the constructions used by Tulenbaev in his proof of~\cite[Proposition~4.3]{Tu83}. This may also help to explain the intuition behind our definition of $\overline{V}$. Recall that Tulenbaev constructs an action of $\St(\Phi, B)$ (of course, only for $\Phi = \rA_\ell$) on a different set, which we denote by $\overline{V}_T$. This set $\overline{V}_T$ is defined as the quotient of the set of triples $V_T = \overline{G}_M^{\geq 0} \times \St(\Phi, A[X\inv]) \times (1+M)^\times$ modulo the equivalence relation given by $(p \gamma, h, u) \sim (p, \gamma h, u)$, $\gamma \in G^0_M$. Tulenbaev denotes the image of $(p, h, u)$ in $\overline{V}_T$ by $[p, h, u]$.

The isomorphism between the sets $\overline{V}$ and $\overline{V}_T$ is given by the following two maps:
\begin{equation} \label{eq:VVT} \xymatrix @R0.3pc {\overline{V} \ar[r]^{\cong} & \ar[l] \overline{V}_T \\ (g, [h], u) \ar@{|->}[r] & [p(g, h, u), h^{-1}, u] \\ (p \cdot j_-(h) \cdot \{u, X\}, [h^{-1}], u) & \ar@{|->}[l] [p, h, u]. } \end{equation}
It is easy to see that these two maps are mutually inverse. This isomorphism allows us to treat these two sets as more or less the same object. In certain situations, however, it is preferable to work with one of them rather than the other. For example, specifying the action of $\St(\Phi, B)$ in terms of $\overline{V}$ leads to much shorter calculations in Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1}, while the statements of~\cref{sec:P1glueing} look more natural when stated for~$\overline{V}_T$.

Now we are ready to proceed with the construction of the action of $\St(\Phi, B)$ on $\overline{V}$.

@@

We are ready to formulate the main result of this subsection.
\begin{prop} \label{prop43}
For $\Phi$ as in the statement of~\cref{lemma33}.
the operators $T_\alpha$ defined above specify a well-defined action of $\St(\Phi, B)$ on $\overline{V} \cong \overline{V}$.

This action satisfies the following additional property. For any $h_1 \in \St(\Phi, A[X\inv])$ holds $j_B^-(h_1) \cdot [1, h, u] = [1, h_1 h, u]$, where $j_B^-$ denotes the map $\St(\Phi, A[X\inv]) \to \St(\Phi, B)$ (here we identify $\overline{V}$ with $\overline{V}_T$ using the isomorphism~\eqref{eq:VVT}).
\end{prop}
\begin{proof}
  By~\cref{superfluous-relations} and Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1} the action of $\St^{\leq 1}(\Phi, B)$ on $\overline{V}$ given by~\eqref{T_1} and~\eqref{T_leq0} is well-defined. On the other hand, the group $\St^{\leq 1}(\Phi, B)$ is isomorphic to $\St(\Phi, B)$ by~\cref{lemma33} (set $t = X^{-1}$).

@@

\subsection{Proof of the $\mathbb{P}^1$-glueing theorem.} \label{sec:P1glueing}
We retain notation of the previous subsection and assume additionally that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, R)$ is injective. 
This assumption allows us to consider $\St(\Phi, B)$ as a subgroup of $\St(\Phi, R)$.
By~\cref{lem:lemma32} $\St(\Phi, B)$ contains $\overline{G}^{\geq 0}_M$.
%We denote by $t$ the map $\St(\Phi, R, I) \to \St(\Phi, B)$ from.
\begin{lemma}\label{lem:action} For any $[p, h, u]\in \overline{V}_T$ and any $p_1 \in \overline{G}^{\geq 0}_M$ holds $p_1 \cdot [p, h, u] = [p_1p, h, u]$. \end{lemma}
\begin{proof} First of all, notice that the assertion of the lemma can be verified directly in the case $p_1 = z_\alpha(f, a)$ for all $a\in A$, $f\in M[X]$. Using~\cref{SR:additivity} and the fact that $x_\beta(X^2f) \in P_\alpha(0)$ for all $\beta \in \Phi$ it is not hard to show that the assertion also holds for $p_1 = z_\alpha(X^2f, aX^{-1})$.

It suffices to verify the assertion of the lemma for the generators $p_1 = z_\alpha(f, \xi)$ of $\overline{G}_M^{\geq 0}$ for all $f \in M[X]$, $\xi\in A[X]$, $\alpha \in \Phi$. We accomplish this by 
induction on the degree of $\xi$ in $X$. The base of induction is clear.

Suppose that the assertion holds for all $p_1 = z_\alpha(f, \xi)$ for which $\xi$ has degree $\leq n$.
Substituting $s := fX$, $\eta := X^{-1}$, $\xi := X\xi$ in~\eqref{Z5} we obtain the following equality in $\St(\Phi, B)$:
\begin{multline*}
 z_{\alpha+\beta}(f, X\xi) = x_\alpha(\epsilon Xf) \cdot x_{-\beta}(-faX^2) \cdot x_{\beta}(af) \cdot x_{\alpha+\beta}(f) \cdot \\ \cdot z_\alpha(-\epsilon Xf, -\epsilon \xi) \cdot x_{-\alpha}(-\epsilon \xi^2fX) \cdot x_{-\alpha-\beta}(- f \xi^2X^2) \cdot z_{-\beta}(X^2f\xi, -X^{-1}),\text{ where $\epsilon = N_{\alpha, \beta}$.}
\end{multline*}
From the first part of the proof and inductive assumption we obtain that the assertion of the lemma holds for all the factors in the right-hand side and, therefore, also holds holds for $p_1 = z_\alpha(f, X\xi)$. It is easy to deduce from this that the assertion also holds for $p_1 = z_\alpha(f, X\xi + a) = x_\alpha(-a) \cdot z_\alpha(f, X\xi) \cdot x_\alpha(a)$. 
\end{proof}

\begin{theorem}
 Let $\Phi$ be a root system of type $\rA_{\geq 4}, \rD_{\geq 5}$ or $\rE_{6,7,8}$.
 %Suppose that the map $j_R \colon \St(\Phi, B) \to \St(\Phi, A[X, X\inv])$ is injective.  
 Then the map $j_-$ is injective and the following commutative square is pullback
 \begin{equation} \nonumber
  \xymatrix{ \St(\Phi, A) \ar[r] \ar[d] & \St(\Phi, A[X]) \ar[d]^{j_+} \\ \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X\inv]).}
 \end{equation}
\end{theorem}
\begin{proof} Since $j_- = j_R j_B^-$, where $j_B^-$ is as in~\cref{prop43}, it suffices to show that $j_B$ is injective.
 Set $v_0 = [1, 1, 1] \in \overline{V}$ and suppose $g \in \Ker(j_B)$.
 By~\cref{prop43} holds $v_0 = j_B(g) \cdot v_0 = [1, g, 1]$ therefore $g \in G_M^0$ and consequently $g = 1$.

 Now suppose $g_+ \in \St(\Phi, A[X])$ and $g_- \in \St(\Phi, A[X\inv])$ are such that $j_-(g_-) = j_+(g_-)$. From~\cref{field-injectivity} it follows that the image $j_+(g_+) = j_-(g_-)$ in $\St(\Phi, k[X, X\inv])$ belongs to $\St(\Phi, k)$ and therefore coincides with the image in $\St(\Phi, k)$ of some $g_0 \in \St(\Phi, A)$.

 Set $h_+ = g_+g_0^{-1}$ and $h_- = g_-g_0^{-1}$, $v_0 = (1, [j_-(h_-)], 1)$.
 It is clear that $h_+ \in \overline{\St}(\Phi, A[X], M[X])$, $h_- \in \overline{\St}(\Phi, A[X\inv], M[X\inv])$ and that $j_-(h_-) = j_+(h_+)$.
 From~\cref{prop43} and~\cref{lem:action} we obtain that
 \[ [j_-(h_+), h_-, 1] = j_B^+(h_+) \cdot [1, h_-, 1] = j_B^-(h_-) j_B^+(h_+) \cdot [1, 1, 1] = [1, 1, 1], \]
 from which it follows that $h_-$ (and hence $h_+$) come from some element of $G_M^0$.
 \end{proof}
